\phantomsection
\section*{参考文献}
\addcontentsline{toc}{section}{参考文献}

\hypertarget{ref:1}{[1]} Y. Sheng et al., "FlexGen: High-throughput generative inference of large language models with a single GPU," in Proc. Int. Conf. Machine Learning (ICML), 2023.


\hypertarget{ref:2}{[2]} A. Borzunov et al., "Petals: Collaborative inference and fine-tuning of large models," in Proc. 61st Annu. Meeting Assoc. Comput. Linguistics (ACL), vol. 3, 2023, pp. 558-568.


\hypertarget{ref:3}{[3]} Y. Jiang et al., "HexGen: Generative inference of large language model over heterogeneous environment," in Proc. Int. Conf. Machine Learning (ICML), 2024, pp. 21946-21961.


\hypertarget{ref:4}{[4]} G.-I. Yu et al., "Orca: A distributed serving system for Transformer-based generative models," in Proc. USENIX Symp. Operating Systems Design and Implementation (OSDI), 2022, pp. 521-538.


\hypertarget{ref:5}{[5]} W. Kwon et al., "Efficient memory management for large language model serving with PagedAttention," in Proc. ACM Symp. Operating Systems Principles (SOSP), 2023, pp. 611-626.


\hypertarget{ref:6}{[6]} A. Agrawal et al., "Taming throughput-latency tradeoff in LLM inference with Sarathi-Serve," in Proc. USENIX Symp. Operating Systems Design and Implementation (OSDI), 2024, pp. 117-134.


\hypertarget{ref:7}{[7]} Y. Zhong et al., "DistServe: Disaggregating prefill and decoding for goodput-optimized large language model serving," in Proc. USENIX Symp. Operating Systems Design and Implementation (OSDI), 2024.


\hypertarget{ref:8}{[8]} L. Zheng et al., "SGLang: Efficient execution of structured language model programs," in Proc. Adv. Neural Inf. Process. Syst. (NeurIPS), 2024.


\hypertarget{ref:9}{[9]} NVIDIA, "TensorRT-LLM documentation," 2026. [Online]. Available: \href{https://nvidia.github.io/TensorRT-LLM/}{\url{https://nvidia.github.io/TensorRT-LLM/}}


\hypertarget{ref:10}{[10]} NVIDIA, "Triton Inference Server: Dynamic batcher," 2026. [Online]. Available: \href{https://docs.nvidia.com/deeplearning/triton-inference-server/}{\url{https://docs.nvidia.com/deeplearning/triton-inference-server/}}


\hypertarget{ref:11}{[11]} Hugging Face, "Text Generation Inference documentation," 2026. [Online]. Available: \href{https://huggingface.co/docs/text-generation-inference/}{\url{https://huggingface.co/docs/text-generation-inference/}}


\hypertarget{ref:12}{[12]} Y. Jiang, R. Yan, and B. Yuan, "HexGen-2: Disaggregated generative inference of LLMs in heterogeneous environment," in Proc. Int. Conf. Learn. Representations (ICLR), 2025.


\hypertarget{ref:13}{[13]} S. Duan, D. Wang, J. Ren, F. Lyu, Y. Zhang, H. Wu, and X. Shen, "Distributed artificial intelligence empowered by end-edge-cloud computing: A survey," IEEE Communications Surveys \& Tutorials, vol. 25, no. 1, pp. 591-624, 2023.


\hypertarget{ref:14}{[14]} Y. Huang et al., "GPipe: Efficient training of giant neural networks using pipeline parallelism," in Proc. Adv. Neural Inf. Process. Syst. (NeurIPS), 2019, pp. 103-112.


\hypertarget{ref:15}{[15]} D. Narayanan et al., "Efficient large-scale language model training on GPU clusters using Megatron-LM," in Proc. Int. Conf. High Performance Computing, Networking, Storage and Analysis (SC), 2021.


\hypertarget{ref:16}{[16]} L. Zheng et al., "Alpa: Automating inter- and intra-operator parallelism for distributed deep learning," in Proc. USENIX Symp. Operating Systems Design and Implementation (OSDI), 2022, pp. 559-578.


\hypertarget{ref:17}{[17]} R. Y. Aminabadi et al., "DeepSpeed-Inference: Enabling efficient inference of transformer models at unprecedented scale," in Proc. Int. Conf. High Performance Computing, Networking, Storage and Analysis (SC), 2022.


\hypertarget{ref:18}{[18]} P. Patel et al., "Splitwise: Efficient generative LLM inference using phase splitting," in Proc. ACM/IEEE Int. Symp. Computer Architecture (ISCA), 2024, pp. 118-132.


\hypertarget{ref:19}{[19]} B. Sun et al., "Llumnix: Dynamic scheduling for large language model serving," in Proc. USENIX Symp. Operating Systems Design and Implementation (OSDI), 2024, pp. 173-191.


\hypertarget{ref:20}{[20]} R. Prabhu, A. Nayak, J. Mohan, R. Ramjee, and A. Panwar, "vAttention: Dynamic memory management for serving LLMs without PagedAttention," in Proc. ACM Int. Conf. Architectural Support for Programming Languages and Operating Systems (ASPLOS), 2025, pp. 1133-1150.


\hypertarget{ref:21}{[21]} R. Qin et al., "Mooncake: Trading more storage for less computation -- A KVCache-centric architecture for serving LLM chatbot," in Proc. USENIX Conf. File and Storage Technologies (FAST), 2025, pp. 155-170.


\hypertarget{ref:22}{[22]} Y. Liu et al., "CacheGen: KV cache compression and streaming for fast large language model serving," in Proc. ACM SIGCOMM, 2024, pp. 38-56.


\hypertarget{ref:23}{[23]} V. Srivatsa et al., "Preble: Efficient distributed prompt scheduling for LLM serving," in Proc. Int. Conf. Learn. Representations (ICLR), 2025.


\hypertarget{ref:24}{[24]} B. Gao et al., "Cost-efficient large language model serving for multi-turn conversations with CachedAttention," in Proc. USENIX Annu. Tech. Conf. (ATC), 2024, pp. 111-126.


\hypertarget{ref:25}{[25]} T. Ouyang, Z. Zhou, and X. Chen, "Follow Me at the Edge: Mobility-aware dynamic service placement for mobile edge computing," IEEE J. Sel. Areas Commun., vol. 36, no. 10, pp. 2333-2345, Oct. 2018.


\hypertarget{ref:26}{[26]} Y. Sheng et al., "Fairness in serving large language models," in Proc. USENIX Symp. Operating Systems Design and Implementation (OSDI), 2024, pp. 965-988.



\hypertarget{ref:27}{[27]} S. Ye, B. Ouyang, L. Zeng, T. Qian, X. Chu, J. Tang, and X. Chen, "Jupiter: Fast and Resource-Efficient Collaborative Inference of Generative LLMs on Edge Devices," in Proc. IEEE INFOCOM, 2025.

\hypertarget{ref:28}{[28]} B. Zhang, J. Zhang, J. Hou, and Y. Wang, "TensAllo: Adaptive Deployment of LLMs on Resource-Constrained Heterogeneous Edge Devices," in Proc. IEEE INFOCOM, 2025.

\hypertarget{ref:29}{[29]} M. Zhang, J. Cao, X. Shen, and Z. Cui, "EdgeShard: Efficient LLM Inference via Collaborative Edge Computing," IEEE Internet of Things Journal, 2025.

\hypertarget{ref:30}{[30]} S. Ye et al., "Resource-Efficient Collaborative Edge Transformer Inference With Hybrid Model Parallelism," IEEE Transactions on Mobile Computing, vol. 24, no. 10, pp. 10945-10962, 2025.

\hypertarget{ref:31}{[31]} J. Du, Y. Wei, S. Ye, J. Jiang, X. Chen, D. Huang, and Y. Lu, "Co-Designing Transformer Architectures for Distributed Inference With Low Communication," IEEE Transactions on Parallel and Distributed Systems, vol. 36, no. 4, 2025.

\hypertarget{ref:32}{[32]} X. Feng et al., "E2LLM: Structure-Guided Efficient Inference for LLMs in Distributed Edge-IoT Environments," IEEE Transactions on Mobile Computing, 2026.

\hypertarget{ref:33}{[33]} Y. Lin, S. Peng, C. Lu, C. Xu, and K. Ye, "FlexPipe: Adapting Dynamic LLM Serving Through Inflight Pipeline Refactoring in Fragmented Serverless Clusters," in Proc. EuroSys, 2026.

\hypertarget{ref:34}{[34]} G. Qu, T. Li, Q. Chen, X. Chen, and S. Zhou, "SLIDE: Simultaneous Model Downloading and Inference at the Wireless Network Edge," IEEE Transactions on Mobile Computing, 2026.

\hypertarget{ref:35}{[35]} D. Xu et al., "EdgeLLM: Fast On-Device LLM Inference With Speculative Decoding," IEEE Transactions on Mobile Computing, vol. 24, no. 4, pp. 3256-3271, 2025.

\hypertarget{ref:36}{[36]} F. Chen, P. Li, T. H. Luan, Z. Su, and J. Deng, "SPIN: Accelerating Large Language Model Inference with Heterogeneous Speculative Models," in Proc. IEEE INFOCOM, 2025.

\hypertarget{ref:37}{[37]} X. Miao et al., "SpecInfer: Accelerating Large Language Model Serving with Tree-based Speculative Inference and Verification," in Proc. ACM Int. Conf. Architectural Support for Programming Languages and Operating Systems (ASPLOS), 2024.

\hypertarget{ref:38}{[38]} R. Kong et al., "SwapMoE: Serving Off-the-shelf MoE-based Large Language Models with Tunable Memory Budget," in Proc. Annu. Meeting Assoc. Comput. Linguistics (ACL), 2024.

\hypertarget{ref:39}{[39]} H. Yu, X. Cui, H. Zhang, H. Wang, and H. Wang, "Taming Latency-Memory Trade-Off in MoE-Based LLM Serving via Fine-Grained Expert Offloading," in Proc. EuroSys, 2026.

\hypertarget{ref:40}{[40]} L. Xue, Y. Fu, Z. Lu, C. Sun, L. Mai, and M. Marina, "MoE-Infinity: Efficient MoE Inference on Personal Machines with Sparsity-Aware Expert Cache," in Proc. Int. Conf. Machine Learning (ICML), 2025.

\hypertarget{ref:41}{[41]} R. Hwang et al., "Pre-gated MoE: An Algorithm-System Co-Design for Fast and Scalable MoE Inference," in Proc. ACM/IEEE Int. Symp. Computer Architecture (ISCA), 2024.

\hypertarget{ref:42}{[42]} R. Yi, L. Guo, S. Wei, A. Zhou, S. Wang, and M. Xu, "EdgeMoE: Empowering Sparse Large Language Models on Mobile Devices," IEEE Transactions on Mobile Computing, 2025.

\hypertarget{ref:43}{[43]} N. Xue et al., "WDMoE: Wireless Distributed Mixture of Experts for Large Language Models," IEEE Transactions on Wireless Communications, 2025.

\hypertarget{ref:44}{[44]} R. Lu, Y. Liu, and D. Wang, "P4LLM: Protecting Embedding Privacy against Model Inversion Attacks for EaaS LLM Serving via Token-wise Partition and Perturbation," IEEE Transactions on Mobile Computing, 2026.

\hypertarget{ref:45}{[45]} W. Zhang, Z. Wu, Y. Mu, R. Ning, B. Liu, N. Sarda, M. Lee, and F. Lai, "JITServe: SLO-aware LLM Serving with Imprecise Request Information," in Proc. USENIX Symp. Networked Systems Design and Implementation (NSDI), 2026.

\hypertarget{ref:46}{[46]} K. Hong et al., "SOLA: Optimizing SLO Attainment for Large Language Model Serving with State-Aware Scheduling," in Proc. Conf. Machine Learning and Systems (MLSys), 2025.

\hypertarget{ref:47}{[47]} A. Patke et al., "One Queue Is All You Need: Resolving Head-of-Line Blocking in Large Language Model Serving," in Proc. Workshop on Cloud Intelligence/AIOps (AIOps@ASPLOS), 2024.

\hypertarget{ref:48}{[48]} N. Ling, G. Chen, and L. Zhong, "TimelyLLM: Segmented LLM Serving System for Time-sensitive Robotic Applications," in Proc. ACM MobiSys, 2026.

\hypertarget{ref:49}{[49]} B. Wu et al., "FastServe: Iteration-Level Preemptive Scheduling for Large Language Model Inference," in Proc. USENIX Symp. Networked Systems Design and Implementation (NSDI), 2026.

\hypertarget{ref:50}{[50]} J. Chen et al., "TokenFlow: Responsive LLM Text Streaming Serving under Request Burst via Preemptive Scheduling," in Proc. EuroSys, 2026.

\hypertarget{ref:51}{[51]} K. Zhu et al., "NanoFlow: Towards Optimal Large Language Model Serving Throughput," in Proc. USENIX Symp. Operating Systems Design and Implementation (OSDI), 2025.

\hypertarget{ref:52}{[52]} C. Ruan et al., "Libra: Flexible Request Partitioning and Scheduling for Serving Unbalanced and Dynamic LLM Workloads," in Proc. USENIX Symp. Networked Systems Design and Implementation (NSDI), 2026.

\hypertarget{ref:53}{[53]} Y. Shen, Y. Zhang, and D. Yuan, "A Hybrid Online and Offline Requests Inference Serving System for LLM in Private Computer Environment," IEEE Transactions on Computers, vol. 75, no. 3, 2026.

\hypertarget{ref:54}{[54]} Z. Li et al., "AdaServe: Accelerating Multi-SLO LLM Serving with SLO-Customized Speculative Decoding," in Proc. EuroSys, 2026.

\hypertarget{ref:55}{[55]} J. Yang, Q. Wu, Z. Feng, Z. Zhou, D. Guo, and X. Chen, "Quality-of-Service Aware LLM Routing for Edge Computing with Multiple Experts," IEEE Transactions on Mobile Computing, 2025.

\hypertarget{ref:56}{[56]} Y. Li, J. Guo, Z. Tang, X. Ding, J. Wang, T. Wang, and W. Jia, "Cloud-Edge System for Scheduling Unpredictable LLM Requests With Combinatorial Bandit," IEEE Transactions on Services Computing, vol. 18, no. 6, pp. 3567-3580, 2025.

\hypertarget{ref:57}{[57]} I. Ong et al., "RouteLLM: Learning to Route LLMs from Preference Data," in Proc. Int. Conf. Learn. Representations (ICLR), 2025.

\hypertarget{ref:58}{[58]} F. Wu and S. Silwal, "PORT: Efficient Training-Free Online Routing for High-Volume Multi-LLM Serving," in Proc. Adv. Neural Inf. Process. Syst. (NeurIPS), 2025.

\hypertarget{ref:59}{[59]} Y. He, J. Fang, F. R. Yu, and V. C. Leung, "Large Language Models (LLMs) Inference Offloading and Resource Allocation in Cloud-Edge Computing: An Active Inference Approach," IEEE Transactions on Mobile Computing, vol. 23, no. 12, pp. 11253-11264, 2024.

\hypertarget{ref:60}{[60]} X. Xu, G. Feng, Y. Liu, S. Qin, J. Wang, and Y. Wang, "Joint Inference Offloading and Model Caching for Small and Large Language Model Collaboration," IEEE Transactions on Mobile Computing, vol. 25, no. 2, 2026.

\hypertarget{ref:61}{[61]} A. Ghosh, N. Sharma, S. Mishra, R. Misra, and S. K. Das, "MGCO: Mobility-Aware Generative Computation Offloading in Edge-Cloud Systems," IEEE Transactions on Services Computing, vol. 19, no. 1, pp. 477-490, 2026.

\hypertarget{ref:62}{[62]} Y. Yue et al., "MasRouter: Learning to Route LLMs for Multi-Agent Systems," in Proc. Annu. Meeting Assoc. Comput. Linguistics (ACL), 2025.

\hypertarget{ref:63}{[63]} S. Nielsen, E. Cetin, P. Schwendeman, Q. Sun, J. Xu, and Y. Tang, "Learning to Orchestrate Agents in Natural Language with the Conductor," in Proc. Int. Conf. Learn. Representations (ICLR), 2026.

\hypertarget{ref:64}{[64]} J. Zhang et al., "AFlow: Automating Agentic Workflow Generation," in Proc. Int. Conf. Learn. Representations (ICLR), 2025.

\hypertarget{ref:65}{[65]} X. Chen, J. Cao, Y. Sahni, M. Zhang, Z. Liang, and L. Yang, "Mobility-Aware Dependent Task Offloading in Edge Computing: A Digital Twin-Assisted Reinforcement Learning Approach," IEEE Transactions on Mobile Computing, vol. 24, no. 4, pp. 2979-2994, 2025.

\hypertarget{ref:66}{[66]} J. Jiang et al., "Efficient KV Cache Spillover Management on Memory-Constrained GPU for LLM Inference," IEEE Transactions on Parallel and Distributed Systems, vol. 37, no. 1, 2026.

\hypertarget{ref:67}{[67]} S. Gao, Q. Wang, S. Zeng, Y. Lu, and J. Shu, "Weaver: Efficient Multi-LLM Serving with Attention Offloading," in Proc. USENIX Annu. Tech. Conf. (ATC), 2025, pp. 587-595.

\hypertarget{ref:68}{[68]} M. Xu, D. Niyato, and C. G. Brinton, "Serving Long-Context LLMs at the Mobile Edge: Test-Time Reinforcement Learning-Based Model Caching and Inference Offloading," IEEE/ACM Transactions on Networking, vol. 34, pp. 3808-3823, 2026.

\hypertarget{ref:69}{[69]} J. Wang et al., "Context-Aware AIGC Service Migration in Edge Intelligence Networks via Transformer DRL," IEEE Transactions on Services Computing, vol. 19, no. 2, pp. 1020-1033, 2026.


\hypertarget{ref:70}{[70]} LMDeploy Team, "LMDeploy documentation," 2026. [Online]. Available: \href{https://lmdeploy.readthedocs.io/}{\url{https://lmdeploy.readthedocs.io/}}

\hypertarget{ref:71}{[71]} A. Du, J. Jia, S. Dustdar, J. Chen, and X. Wang, "Online Service Placement, Task Scheduling, and Resource Allocation in Hierarchical Collaborative MEC Systems," IEEE Transactions on Services Computing, vol. 18, no. 2, pp. 983-997, 2025.

\hypertarget{ref:72}{[72]} X. Zhuang, J. Wu, H. Wu, T. Zhang, and L. Gao, "Joint Optimization of Model Inferencing and Task Offloading for MEC-Empowered Large Vision Model Services," in Proc. IEEE INFOCOM, 2025.

\hypertarget{ref:73}{[73]} Z. Zhou et al., "A Survey on Efficient Inference for Large Language Models," arXiv:2404.14294, 2024.

\hypertarget{ref:74}{[74]} R. Zhen et al., "Taming the Titans: A Survey of Efficient LLM Inference Serving," in Proc. Int. Natural Language Generation Conf. (INLG), 2025, pp. 522-541.

\hypertarget{ref:75}{[75]} J. Pan and G. Li, "A Survey of LLM Inference Systems," arXiv:2506.21901, 2025.






\hypertarget{ref:76}{[76]} 华为，《边缘 AI 计算技术合作项目资料》，2026.

\hypertarget{ref:77}{[77]} vLLM Ascend Team, "vLLM Ascend documentation," 2026. [Online]. Available: \href{https://docs.vllm.ai/projects/ascend/en/latest/}{\url{https://docs.vllm.ai/projects/ascend/en/latest/}}

\hypertarget{ref:78}{[78]} vLLM Ascend Team, "vLLM Ascend installation guide," 2026. [Online]. Available: \href{https://docs.vllm.ai/projects/ascend/en/latest/installation.html}{\url{https://docs.vllm.ai/projects/ascend/en/latest/installation.html}}

\hypertarget{ref:79}{[79]} vLLM Ascend Team, "vLLM Ascend features," 2026. [Online]. Available: \href{https://docs.vllm.ai/projects/ascend/en/latest/features/}{\url{https://docs.vllm.ai/projects/ascend/en/latest/features/}}

\hypertarget{ref:80}{[80]} vLLM Ascend Team, "KV Cache Pool," 2026. [Online]. Available: \href{https://docs.vllm.ai/projects/ascend/en/latest/features/kv_cache_pool.html}{\url{https://docs.vllm.ai/projects/ascend/en/latest/features/kv_cache_pool.html}}

\hypertarget{ref:81}{[81]} vLLM Ascend Team, "Speculative decoding," 2026. [Online]. Available: \href{https://docs.vllm.ai/projects/ascend/en/latest/features/spec_decode.html}{\url{https://docs.vllm.ai/projects/ascend/en/latest/features/spec_decode.html}}

\hypertarget{ref:82}{[82]} Huawei Ascend, "Model inference," 2026. [Online]. Available: \href{https://www.hiascend.com/software/model-inference}{\url{https://www.hiascend.com/software/model-inference}}

\hypertarget{ref:83}{[83]} vLLM Project, "OpenAI-compatible server," 2026. [Online]. Available: \href{https://docs.vllm.ai/en/latest/serving/openai_compatible_server.html}{\url{https://docs.vllm.ai/en/latest/serving/openai_compatible_server.html}}


